
---
title: "Auto-mpg"
author: "Fraydie Kaplan"
date: "2023-12-07"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


## Linear Regression Analysis

```{r setup, include=FALSE}
# Set a CRAN mirror
options(repos = c(CRAN = "https://cran.rstudio.com"))

# Install/load necessary packages
install.packages("caret")
library(tidyverse)
library(caret)
```
## Load and Explore the Dataset
```{r}
# Load the dataset
df <- read.csv("C:/Users/Fraydie/Documents/random college/Semester 4/Data Science/Final Project/auto-mpg.csv")

# Display summary
summary(df)
str(df)

# Change horsepower to numeric
df$horsepower <- as.numeric(df$horsepower)

# Confirm horsepower was changed to numeric
str(df)
```

# Exploratory Data Analysis

## Scatter Plot of MPG vs. Weight
```{r}
# Scatter plot of mpg vs. weight
ggplot(df, aes(x = weight, y = mpg)) +
  geom_point() +
  labs(title = "Scatter Plot of MPG vs. Weight",
       x = "Weight",
       y = "MPG") +
  theme_minimal()
```
## Pair Plot to Visualize Relationships
```{r}
# Pair plot to visualize relationships between variables
features_to_plot <- c("mpg", "displacement", "horsepower", "weight", "acceleration")
pairs(df[, features_to_plot])
```
## Box Plot for MPG and Number of Cylinders
```{r}
# Box plot to show the distribution of mpg for different numbers of cylinders
ggplot(df, aes(x = as.factor(cylinder), y = mpg)) +
  geom_boxplot() +
  labs(title = "Box Plot of MPG for Different Numbers of Cylinders",
       x = "Number of Cylinders",
       y = "MPG") +
  theme_minimal()
```
# Linear Regression Modeling

## Simple Linear Regression
```{r}
# Simple Linear Regression (using the first 300 samples)
model_simple <- lm(mpg ~ weight, data = head(df, 300))
summary(model_simple)
```
## Multiple Linear Regression
```{r}
# Multiple Linear Regression
features <- c("cylinder", "displacement", "horsepower", "weight", "acceleration", "model.year", "origin")
model_multiple <- lm(mpg ~ ., data = head(df, 300)[, c('mpg', features)])
summary(model_multiple)
```
## Clean Data for Multiple Regression
```{r}
# Clean data for multiple regression (excluding "car.name")
df_cleaned <- df[1:300, c("mpg", features)]
df_cleaned <- na.omit(df_cleaned)
# Ensure "car.name" is not included in remaining_data
remaining_data <- df[301:398, c("mpg", features)]
remaining_data$horsepower <- as.numeric(remaining_data$horsepower)
str(remaining_data)
```
## Regression Model Logs
```{r}
# Log of Regression Models
log_simple <- list("Simple Linear Regression",
                   "Multiple R-squared" = summary(model_simple)$r.squared,
                   "Adjusted R-squared" = summary(model_simple)$adj.r.squared,
                   "Complete Linear Regression equation" = paste(names(coef(model_simple)), ": ", coef(model_simple), collapse = ", "))

log_multiple <- list("Multiple Linear Regression",
                     "Multiple R-squared" = summary(model_multiple)$r.squared,
                     "Adjusted R-squared" = summary(model_multiple)$adj.r.squared,
                     "Complete Linear Regression equation" = paste(names(coef(model_multiple)), ": ", coef(model_multiple), collapse = ", "))

# Log the regression models
log_simple <- unlist(log_simple)
log_multiple <- unlist(log_multiple)
log_simple
log_multiple
```
# Diagnostic Plots for Regression Models

## Simple Linear Regression
```{r}
# Diagnostic plots for Simple Linear Regression
plot(model_simple, which = c(1, 2, 3, 5))
```
## Multiple Linear Regression
```{r}
# Diagnostic plots for Multiple Linear Regression
plot(model_multiple, which = c(1, 2, 3, 5))
```
# Predictions and Residual Analysis

## Predictions for the Remaining 98 Samples
```{r}
# Predictions for the Remaining 98 Samples
predictions_simple <- predict(model_simple, newdata = remaining_data)
predictions_multiple <- predict(model_multiple, newdata = remaining_data[, c('mpg', features)])
```
## Residual Plot for Simple Linear Regression
```{r}
# Residual Plot for Simple Linear Regression
residuals_simple <- remaining_data$mpg - predictions_simple
plot(predictions_simple, residuals_simple, main = "Residual Plot - Simple Linear Regression", xlab = "Predicted MPG", ylab = "Residuals")
```
## Histogram of Residuals for Simple Linear Regression
```{r}
# Histogram of Residuals for Simple Linear Regression
hist(residuals_simple, main = "Histogram of Residuals - Simple Linear Regression", xlab = "Residuals", ylab = "Frequency")
```
## Residual Plot for Multiple Linear Regression
```{r}
# Residual Plot for Multiple Linear Regression
residuals_multiple <- remaining_data$mpg - predictions_multiple
plot(predictions_multiple, residuals_multiple, main = "Residual Plot - Multiple Linear Regression", xlab = "Predicted MPG", ylab = "Residuals")
```
## Histogram of Residuals for Multiple Linear Regression
```{r}
# Histogram of Residuals for Multiple Linear Regression
hist(residuals_multiple, main = "Histogram of Residuals - Multiple Linear Regression", xlab = "Residuals", ylab = "Frequency")
```





